A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-08 paper.Rmd
section 5 intro finished

## Statistical significance

@@

Data collection resulted in a total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` for the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` for the heteroskedasticity model. This was done in three collection periods, called I, II, III, with `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluated by 160, 150 and 133 participants, respectively. There were `r nrow(filter(vi_survey, attention_check))` attention check lineups used to filter false participant attempts. To estimate $\alpha$ for calculating statistical significance there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of null lineups. Neither the attention checks nor null lineups were used for the subsequent analysis. The de-identified data, `vi_survey`,  is made available in the R package, `visage`.

For clarity and consistency, the analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only on results from lineups generated with uniform fitted values, a total of `r unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id))` lineups and `nrow(filter(vi_survey, !attention_check, !null_lineup))` evaluations. More evaluations were made using the uniform fitted value distribution, and it avoids potentially confounding results of the analysis by combining values from different power curves. We also expect that visual patterns are more clearly visible when the fitted value distribution is uniform. The results in Section \ref{effect-of-fitted-value-distributions} examine the effect of fitted value distribution on power.

>>>>

## Statistical significance~\label{sig}

@@

Data collection used a total of `r vi_survey %>% count(unique_lineup_id) %>% nrow()` lineups, and resulted in a total of `r nrow(filter(vi_survey, !attention_check))` evaluations from `r vi_survey %>% count(set, exp) %>% nrow()` participants. Roughly half corresponded to the two models, non-linearity and heteroskedasticiy, and the three collection periods had similar numbers in each. <!-- `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` for the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` for the heteroskedasticity model. This was done in three collection periods (I, II, III), with `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluated by 160, 150 and 133 participants, respectively.--> Each participant received two of the `r vi_survey %>% filter(attention_check) %>% count(unique_lineup_id) %>% nrow()` attention check lineups which were used to filter results of participants who were clearly not making an honest effort (only `r vi_survey %>% filter(attention_check, !detect) %>% count(exp, set) %>% nrow()` of `r vi_survey %>% count(set, exp) %>% nrow()`). To estimate $\alpha$ for calculating statistical significance (see Section \ref{sig}) there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of `r vi_survey %>% filter(null_lineup, !attention_check) %>% count(unique_lineup_id) %>% nrow()` null lineups. Neither the attention checks nor null lineups were used for the subsequent analysis. The de-identified data, `vi_survey`,  is made available in the R package, `visage`.

The data was collected on lineups constructed from four different fitted value distributions: uniform, normal, skewed and discrete. More data was collected on the uniform (each evaluated by 11 participants) than the others (each evaluated by 5 participants). The analysis in Sections \ref{power-analysis}-\ref{hetero-analysis}, uses only results from lineups generated with uniform fitted values, a total <!--of `r length(unique(pull(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"), unique_lineup_id)))` lineups and--> `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations. This was decided in order to compare the conventional and visual test performance for an optimal scenario. Section \ref{effect-of-fitted-value-distributions} examines how the results may be affected if the fitted value distribution was different.

### 2023-04-08 paper.tex
section 5 intro finished

\hypertarget{statistical-significance}{%
\subsection{Statistical significance}\label{statistical-significance}}

@@

Data collection resulted in a total of 7974 lineup evaluations, 3744 for
the non-linearity model, and 4230 for the heteroskedasticity model. This
was done in three collection periods, called I, II, III, with 2880, 2700
and 1674 lineups evaluated by 160, 150 and 133 participants,
respectively. There were 886 attention check lineups used to filter
false participant attempts. To estimate \(\alpha\) for calculating
statistical significance there were 720 evaluations of null lineups.
Neither the attention checks nor null lineups were used for the
subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
made available in the R package, \texttt{visage}.

For clarity and consistency, the analysis in Sections
\ref{power-analysis}-\ref{hetero-analysis}, uses only on results from
lineups generated with uniform fitted values, a total of 4135 lineups
and XXX evaluations. More evaluations were made using the uniform fitted
value distribution, and it avoids potentially confounding results of the
analysis by combining values from different power curves. We also expect
that visual patterns are more clearly visible when the fitted value
distribution is uniform. The results in Section
\ref{effect-of-fitted-value-distributions} examine the effect of fitted
value distribution on power.

@@

\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only). The power curves are estimated using logistic regression, and the horizontal lines of dots represent human evaluations of lineups, with 0, 1 indicating non-detect and detect, respectively.  lines The visual test has multiple power curves estimated from bootstrap samples. The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
\end{figure}

>>>>

\hypertarget{statistical-significance}{%
\subsection{\texorpdfstring{Statistical
significance\textasciitilde{}\label{sig}}{Statistical significance\textasciitilde{}}}\label{statistical-significance}}

@@

Data collection used a total of 1176 lineups, and resulted in a total of
7974 evaluations from 443 participants. Roughly half corresponded to the
two models, non-linearity and heteroskedasticiy, and the three
collection periods had similar numbers in each. Each participant
received two of the 24 attention check lineups which were used to filter
results of participants who were clearly not making an honest effort
(only 4 of 443). To estimate \(\alpha\) for calculating statistical
significance (see Section \ref{sig}) there were 720 evaluations of 36
null lineups. Neither the attention checks nor null lineups were used
for the subsequent analysis. The de-identified data,
\texttt{vi\_survey}, is made available in the R package,
\texttt{visage}.

The data was collected on lineups constructed from four different fitted
value distributions: uniform, normal, skewed and discrete. More data was
collected on the uniform (each evaluated by 11 participants) than the
others (each evaluated by 5 participants). The analysis in Sections
\ref{power-analysis}-\ref{hetero-analysis}, uses only results from
lineups generated with uniform fitted values, a total 3069 lineup
evaluations. This was decided in order to compare the conventional and
visual test performance for an optimal scenario. Section
\ref{effect-of-fitted-value-distributions} examines how the results may
be affected if the fitted value distribution was different.

@@

\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only). The power curves are estimated using logistic regression, and the horizontal lines of dots represent decisions made by visual tests, with 0, 1 indicating non-reject and reject, respectively. The visual test has multiple power curves estimated from bootstrap samples. The row of scatterplots plots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}\label{fig:polypower}
\end{figure}
